---
title: "Titanic"
author: "Matthew Houser"
date: "7/26/2020"
output:
  pdf_document: default
  html_document: default
---


Setup: Convert character features to factors, add column that says "Survived" or "Died" rather than 1 and 0 for graphical reasons. Rename columns for simplicity.
```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(discrim)
library(readr)
library(tidymodels)
titanic <- read.csv("https://web.stanford.edu/class/archive/cs/cs109/cs109.1166/stuff/titanic.csv") %>% 
  mutate_if(is.character, factor) %>% 
  mutate(Survived_text = case_when(Survived == 1 ~ "Survived", Survived == 0 ~ "Died")) 

names(titanic) <- c("Survived", "Pclass", "Name", "Sex", "Age", "SibSp", "ParCh", "Fare", "Survived_text")
```

Inspect the dataset. We have 8 unique features: Survived, and integer indicating whether they survived or died, what passenger class their ticket was, their name, sex, and age, how many siblings or spouses were on board with them, how many parents or children were on board with them, and their ticket fare. 
```{r}
str(titanic)
```
Let's find how many people in the dataset Survived
```{r}
titanic %>% 
  count(Survived_text)
```
Let's facet survival by sex
```{r}
titanic %>% 
  group_by(Sex) %>% 
  count(Survived_text)
```


Exploratory Data Analysis: 
```{r}
titanic %>% 
  group_by(Sex, Pclass, Survived_text) %>% 
  tally() %>% 
  ggplot(aes(Survived_text, n, fill = Survived_text)) +
  geom_col() +
  facet_grid(Sex ~ Pclass) +
  geom_text(aes(label = n, vjust = -0.5, hjust = .5), size = 3) +
  scale_y_continuous(limits = c(0, 310)) +
  xlab("Survival Status") +
  ylab("Sum") +
  ggtitle("Number of Titanic Survivors/Casualties by Passenger Class and Sex")

```


```{r}
titanic %>% 
  group_by(Pclass, Sex, Survived_text) %>% 
  tally() %>% 
  mutate(percent = n/sum(n)) %>% 
  ggplot(aes(Survived_text, percent, fill = Survived_text)) +
  geom_col() +
  facet_grid(Sex ~ Pclass) + 
  geom_text(aes(label = round(percent, 2), vjust = -0.5, hjust = .5), size = 3) +
  scale_y_continuous(limits = c(0, 1.05)) +
  xlab("Survival Status") +
  ylab("Percent")+
  ggtitle("Percentage of Titanic Survivors/Casualties by Passenger Class and Sex")

```

```{r}
titanic %>% 
  ggplot(aes(Survived_text, Age)) +
  geom_jitter(aes(color = factor(ParCh)), width = .2) +
  facet_wrap(Sex ~ Pclass) +
  ggtitle("Survival by Sex, Passenger Class and Parents/Children On Board")
  
```

```{r}
titanic %>% 
  ggplot(aes(Survived_text, Age)) +
  geom_jitter(aes(color = factor(SibSp)), width = .25) +
  facet_wrap(Sex ~ Pclass) +
  ggtitle("Survival by Sex, Passenger Class and Siblings/Spouses On Board")

```
```{r}
#Convert Survived column to a factor in order for classification to work
titanic$Survived <- as.factor(titanic$Survived)

#Set seed for reproducibility, split data into training and testing sets
set.seed(1)
titanic_split <- titanic %>% 
  initial_split(prop = .8, strata = Survived)

train <- training(titanic_split)
test <- testing(titanic_split)
```


```{r, message=FALSE, warning=FALSE}
#Create recipe, model specifications and workflow:
titanic_recipe <- recipe(Survived ~ Sex + Age + SibSp + Pclass + ParCh + Fare, data = train) %>% 
  step_downsample(Survived)

glm_spec <- logistic_reg() %>% 
  set_engine("glm")

rf_spec <- rand_forest() %>% 
  set_engine("ranger") %>% 
  set_mode("classification")

lda_spec <- discrim_linear() %>% 
  set_engine("MASS") %>% 
  translate()

qda_spec <- discrim_regularized(frac_common_cov = 0, frac_identity = 0) %>% 
  set_engine("klaR") %>% 
  translate()
  
titanic_wf <- workflow() %>% 
  add_recipe(titanic_recipe)
```

```{r}
#Fit the logistic model
titanic_glm <- titanic_wf %>% 
  add_model(glm_spec) %>% 
  fit(data = train)

titanic_glm
```

```{r}
#Fit the LDA model
titanic_lda <- titanic_wf %>% 
  add_model(lda_spec) %>% 
  fit(data= train)

titanic_lda
```

```{r}
#Fit the QDA
titanic_qda <- titanic_wf %>% 
  add_model(qda_spec) %>% 
  fit(data = train)

titanic_qda
```

```{r}
#Fit the random forest
titanic_rf <- titanic_wf %>% 
  add_model(rf_spec) %>% 
  fit(data = train)

titanic_rf
```

```{r}
#Use the training model on the test set and then show confusion matrix
results <- test %>% 
  bind_cols(predict(titanic_glm, test) %>% 
              rename(.pred_glm = .pred_class)) %>% 
  bind_cols(predict(titanic_lda, test) %>% 
              rename(.pred_lda = .pred_class)) %>% 
  bind_cols(predict(titanic_qda, test) %>% 
              rename(.pred_qda = .pred_class)) %>% 
  bind_cols(predict(titanic_rf, test) %>% 
              rename(.pred_rf = .pred_class))
 
#Confusion matrix for logistic regression
results %>% 
  conf_mat(truth = Survived, estimate = .pred_glm)
#Confusion matrix for LDA
results %>% 
  conf_mat(truth = Survived, estimate = .pred_lda)
#Confusion matrix for QDA
results %>% 
  conf_mat(truth = Survived, estimate = .pred_qda)
#Confusion matrix for Random Forest
results %>% 
  conf_mat(truth = Survived, estimate = .pred_rf)
```

```{r}
#Find sensitivity for each model 
sens_glm <- sensitivity(results, truth = Survived, estimate = .pred_glm)
sens_lda <- sensitivity(results, truth = Survived, estimate = .pred_lda)
sens_qda <- sensitivity(results, truth = Survived, estimate = .pred_qda)
sens_rf <- sensitivity(results, truth = Survived, estimate = .pred_rf)

c(sens_glm$.estimate, sens_lda$.estimate, sens_qda$.estimate, sens_rf$.estimate)
```

```{r}
#Find specificity for each model 
spec_glm <- specificity(results, truth = Survived, estimate = .pred_glm)
spec_lda <- specificity(results, truth = Survived, estimate = .pred_lda)
spec_qda <- specificity(results, truth = Survived, estimate = .pred_qda)
spec_rf <- specificity(results, truth = Survived, estimate = .pred_rf)

c(spec_glm$.estimate, spec_lda$.estimate, spec_qda$.estimate, spec_rf$.estimate)
```

```{r}
#Find accuracy for each model
acc_glm <- accuracy(results, truth = Survived, estimate = .pred_glm)
acc_lda <- accuracy(results, truth = Survived, estimate = .pred_lda)
acc_qda <- accuracy(results, truth = Survived, estimate = .pred_qda)
acc_rf <- accuracy(results, truth = Survived, estimate = .pred_rf)

c(acc_glm$.estimate, acc_lda$.estimate, acc_qda$.estimate, acc_rf$.estimate)
```
